---
title: "Challenge 3"
author: "Giovanni Librizzi"
format:
  html:
    self-contained: true
    code-tools: true
editor: visual
execute: 
  messsage: false
  echo: true
  error: true
---

Loading in packages

```{r packages}
#| message = FALSE

# code chunk for loading packages and importing the data
library(tidyverse)
hiphop <- read_csv(here::here("Week3", "Lab3", "hiphop.csv"))
hiphop_clean <- hiphop |>
  # sets as factors
  mutate(jayz = as.factor(jayz),
         sex = as.factor(sex),
         ethnic = as.factor(ethnic),
         barkley = as.factor(barkley),
         boondocks = as.factor(boondocks),
         monique = as.factor(monique),
         bieber = as.factor(bieber),
         jayz1 = as.factor(jayz1),
         boondocks1 = as.factor(boondocks1),
         familiarity = as.factor(familiarity),
         popularityScore = as.factor(popularityScore),
         fam1 = as.factor(fam1)
         ) |>
  # replaces 0 with NA 
  mutate(city = case_when(
    city == 0 ~ NA_real_, 
    TRUE ~ city
    )) |>
  mutate(county = case_when(
    county == 0 ~ NA_real_, 
    TRUE ~ county
    )) |>
  mutate(countyBlack = case_when(
    countyBlack == 0 ~ NA_real_, 
    TRUE ~ countyBlack
    )) |>
  # white boolean
  mutate(white = if_else(
    condition = ethnic == "white",
    true = TRUE, 
    false = FALSE
    ))

```

# 5.1 Published Comparisons

```{r}
#| message = FALSE

demographic <- hiphop_clean |>
  distinct(subj, .keep_all = TRUE) |>
  select(sex, 
         white, 
         intl, 
         vocal, 
         classical, 
         folk, 
         rock, 
         country,
         pop, 
         alternative, 
         hiphop
         ) 

# genre higher average 
# male vs female
demographic |>
  group_by(sex) |>
  summarize(
    across(
      .cols = intl:hiphop,
      .fns = mean
      )
    )

# white vs non-white
demographic |>
  group_by(white) |>
  summarize(
    across(
      .cols = intl:hiphop,
      .fns = mean
      )
    )
  
```

Rock appeared to have a much higher distribution of men vs women, with men with a mean of 3.05 artists they reported liking in that genre and women at 1.93. Country was the opposite, with women having a mean of 0.78 while men having 0.39. Vocal did appear to have more men listening than women with men having a mean of 1.56 while women had 0.88.

# 5.2 Study Design Critique - Data Ethics

This study was looking at if non-African American young adults learned African-American English vocabulary through listening to hip-hop. One concern I have with this study is that it has a strange racial context where it seems to try and demonstrate that the main way non-Whites know African American context and vocabulary is through hip-hop, barely focusing on other possible ways and influences African Americans have had on American culture. Another concern I have with the study is with how they used the data to find country music's negative correlation with AAE comprehension and claiming that that personal identity is self-constructed.
