---
title: "Lab 9"
author: "Giovanni Librizzi"
format: 
  html: 
    self-contained: true
    code-tools: true
editor: visual
execute: 
  echo: true
  warning: false
  message: false
  error: true
---

# Revision

I accidentally had my labels for the legends in my plot in the wrong order, so i refactored to allow them to be in the right order. I also didn't put the right y label on my regression equation, which I fixed.

### Setup

```{r}
library(tidyverse)
names_a <- read_csv(here::here("Week9", "Lab", "StateNames_A.csv"))
names_a <- names_a |>
  rename(Sex = Gender)
```

# 3. Summarizing & Visualizing the Number of Allisons

1\.

```{r}
allison_data <- names_a |>
  filter(Name == "Allison") |>
  select(-Year) |>
  group_by(State, Sex) |>
  summarize(n = sum(Count)) |>
  pivot_wider(names_from = Sex,
              values_from = n,
              values_fill = 0) |>
  rename(Male = M, Female = F)

```

2\.

```{r}
allison_f <- names_a |>
  filter(Name == "Allison", 
         Sex == "F") |>
  select(-State) |>
  group_by(Year) |>
  summarize(n = sum(Count))
```

3\.

```{r}
allison_f |>
  ggplot(mapping = aes(x = Year, y = n)) +
  geom_line() +
  geom_point() +
  labs(y = NULL,
       title = "Number of Babies With The Name Allison By Year"
       )
```

# 4. Modeling the Number of Allisons

4.  Fit a linear model with the year as the explanatory variable, and the number of Allisons as the response. Similar to #3, each year should have one observation--the total number of Allisons born that year.

```{r}
allison_lm <- lm(n ~ Year, data = allison_f)
```

5.  Write out the estimated regression equation.

    babies_born_named_allison = 209689.8 - 101\*Year

<!-- -->

6.  Plot the residuals of the model, that is, the actual values minus the predicted values. Comment on the residuals - do you see any patterns?

```{r}
allison_lm |>
  broom::augment() |>
  ggplot(mapping = aes(y = .resid, x = .fitted)) +
  geom_point() +
  labs(x = "Fitted",
       y = "Residuals",
       title = "Residuals of the Allison Model"
       )
```

The residuals start tight and spread out over time on the y-axis, so at the end they are very spread out. They almost seem to trend downward too.

7.  From the spread of this model, we can see the name is very varied in popularity, but does seem to be more unpopular now

# 5. Spelling by State

8\.

```{r}
alan_spelling <- names_a |>
  filter(Name %in% c("Allan", "Alan", "Allen")) |>
  group_by(Name, Year) |>
  summarize(n = sum(Count))
```

```{r}
alan_spelling |>
  ggplot(mapping = aes(x = Year, 
                       y = n, 
                       color = fct_reorder2(Name, Year, n)
                       )) +
  geom_line() +
  geom_point() +
  labs(
    y = NULL,
    title = "Number of Babies with Different Spellings of Allen by Year",
    color = "Name"
  ) 
  
```

9\.

```{r}
allan_state <- names_a |>
  filter(Name %in% c("Allan", "Alan", "Allen"),
         Year == 2000,
         Sex == "M",
         State %in% c("CA", "PA")) |>
  select(-Sex, -Year) |>
  group_by(Name, State) 

allan_state_wide <- allan_state |>
  pivot_wider(names_from = Name,
              values_from = Count,
              values_fill = 0)
```

10. 

```{r}
allan_state |>
  group_by(State) |>
  mutate(prop = round(Count / sum(Count), 4),
         prop = str_c(as.character(prop*100), "%")) |>
  select(-Count) |>
  pivot_wider(names_from = Name,
              values_from = prop,
              values_fill = "0%")
```
