## Step 1: Read in the dataset `ages.csv`.

```{r}
# Code to read in the ages.csv data goes here! 
read_csv("Ages_Data/ages.csv", col_names=TRUE)

```

## Step 2: Read in the dataset `ages_tab.txt`

```{r}
# Code to read in the ages_tab.txt data goes here! 
library(tidyverse)
library(readr)
read_tsv("Ages_Data/ages_tab.txt")
# this file isn't properly set up with tabs so this doesn't work !!!!!!

```

## Step 3: Read in the dataset `ages_mystery.txt`

```{r}
# Code to read in the ages_mystery.txt data goes here! 
library(tidyverse)
library(readr)
read_delim("Ages_Data/ages_mystery.txt", delim = "|")

```

## Step 4: Read in the dataset `ages.xlsx`

```{r}
# Code to read in the ages.xlsx data goes here! 
library(tidyverse)
library(readxl)
ages <- read_excel("Ages_Data/ages.xlsx")

```

## Step 5: Both `read.csv` and `read.table` have an optional argument called `stringsAsFactors`. Figure out what this does.

When false, this makes the strings stay as a character type instead of being converted into the factor data type.

## Step 6: Find a way to use `read_csv()` to read `ages.csv` with the variable "Name" as a factor and "Age" as a character.

```{r}
# Code to read in the ages.csv with "Name" as a factor and "Age" as a character goes here! 
read_csv("Ages_Data/ages.csv", col_types = cols(Name = col_factor(), Age = col_character()))


```

------------------------------------------------------------------------
